---
title: "Week4 - Assignment"
author: Rob Hodde
date: 11/22/15
output: html_document
---

Week 4 Assignment
Your assignment should be published to RPubs.com. Place your script in a public repository on GitHub and submit
a link to the URL here.

Exploratory Data Analysis in R. Choose an interesting dataset and use R graphics to describe the data. You may
use base R graphics, or a graphics package of your choice. You should include at least one example of each of the
following:
. histogram
. boxplot
. scatterplot
Do the graphics provide insight into any relationships in the data?
You may use R code that already exists in a book or a web site as a starting point for your work. You need to cite
your source, and you need to clearly delineate what you were given to start with, and what is your own original
analysis, so that your grade can be based on your own work!
This is an open-ended assignment. Full credit will be given for any reasonable effort, but as with most technical
learning work, the more sustained effort over time that you invest, the more you will get out of it.
If you don't know where to start, there is a list of over 700 R datasets here:
http://vincentarelbundock.github.io/Rdatasets/. You can of course also load data from another source, such as a
.CSV file. You'll want to pick a dataset where it makes sense to use histograms, boxplots, and scatterplots.

Data Sources:

All about Eve: http://eve-marketdata.com/developers/mysql.php

Regions: http://www.eve-search.com/thread/1052168/page/1

Products: http://dl.eve-files.com/media/1202/TypeIDs.txt

Daily Sales Summary:  http://eve-marketdata.com/reports.php?step=DailySales

I have built a mySQL database from the data sources above, and composed two views used below for plotting.  I do not know how to move my database to the cloud so that you can actually run this code.  If this is a problem, please let me know. 



The following histogram shows the extent to which products and commodity prices on the MMORGP "Eve Online - Region "The Forge" have varied over the past 30 days. 


```{r}

 # use the R ODBC package to establish a connection to mySQL database and select records into data frames
 require(RODBC)
 cnString <- "MySQL_ANSI;SERVER=localhost;DATABASE=eve;UID=root;PASSWORD=CUNYRBridge4!;OPTION=3;"  # how to load the db to cloud?
 db <- odbcConnect(cnString)
 
 # download the mySQL view records
 df <- sqlQuery(db,"SELECT * FROM vw_top100_theforge_30days", stringsAsFactors = FALSE)  

 # look at just two fields, so we can plot a histogram
 subdf <- subset(df,Price_Range_Percent < 70, select = c(Total_Spend, Price_Range_Percent))  
 
 # create a histogram that illustrates the distribution of total price range for past 30 days 
 hist(subdf$Price_Range_Percent, main = "Eve Forge - Top 100 Products - Monthly Price Spreads", xlab = "Price Range Percent (Min to Max)")
```



Next we will look at pricing for one of the most popular and somewhat price-volatile products on Eve today - the Gecko heavy drone.


```{r}
 

 # download the mySQL view records
 df <- sqlQuery(db,"SELECT * FROM vw_gecko_theforge_30days", stringsAsFactors = FALSE)  
 
plot(price_average ~ date, data = df)
 
 
```
 
The plot above illustrates that the price has steadily declined over the past 30 days.  We can also see that there isn't a visible relationship between daily order volume (quantity) and price:

```{r}

plot(price_average ~ quantity, data = df)

```


Next we will use the boxplot:

```{r}

boxplot(df$price_average)

```

What's interesting about this boxplot is that it quickly tells the user, "If you purchased a Gecko in November for 78-86 billion, you did alright.  Otherwise, you either paid too much, or got a sweet bargain. 






